Better fusion for filters

### 2014-04-02 2014betterfusionforfilters/section/04-ILP.tex
explain ILP variables
move simplification/opt talk, expand explanation of ILP formulation

\subsubsection{The simple version}
\TODO{Explain this simple version}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (no edge from i to j, but there is a fusion benefit)            \\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (an edge from i to j)            \\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (for all edges from i; thus $c_i=0$ only if $\forall j | (i,j) \in E. x_{ij} = 0$)      \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (fusion-preventing edges from i to j)      \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\emptyset \wedge gen_j=\emptyset$)                 \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\{i'\}$) \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_j=\{j'\}$) \\
\end{tabbing}
It is worth noting that for the edge cases, $x_{ij} = 1 \implies \pi_i < \pi_j$, meaning if $i$ and $j$ are not fused together, $j$ must be scheduled after $i$.

\subsubsection{A slightly smarter version}
We can simplify the version above and remove many constraints and variables by noting that if there is a fusion-preventing edge between $i$ and $j$, $x_{ij} = 1$, so the precedence constraint can be simplified to just $\pi_i < \pi_j$, and the $x_{ij}$ variable removed.
Likewise with clusters of different, incompatible types.
The contraction variable $c_i$ can be removed, if $i$ has any fusion-preventing output edges: any such edges use the array and cannot be fused, thus there is no possibility of the array being contracted away.

@@

>>>>

\subsubsection{The simple version}
\paragraph{Acyclic and precedence-preserving}

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (an edge from $i$ to $j$)            \\
           \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (no edge from $i$ to $j$)            \\
           \> \ldots
\end{tabbing}
Here, we look at every pair of nodes $i$ and $j$.

Whether there is an edge between $i$ and $j$ or not, if the two are fused together, then $x_{ij} = 0$.
If $x_{ij} = 0$, then both cases may be simplified to $0 \le \pi_j - \pi_i \le 0$, which is equivalent to $\pi_i = \pi_j$.

Otherwise, if the two nodes are not fused together, then $x_{ij} = 1$.
If there is an edge between $i$ and $j$, the constraint simplifies to $1 \le \pi_j - \pi_i \le N$, where $N$ is the number of nodes in the graph.
This constraint means that the difference between the two $\pi$s must be at least 1, and less than $N$.
Since there are $N$ nodes, the maximum difference between any two $\pi$s would be at most $N$, so the upper bound of $N$ is large enough to be safely ignored.
This means the constraint can roughly be translated to $\pi_i < \pi_j$, which enforces the acyclic constraint.

If there is no edge between $i$ and $j$, the constraint simplifies to $-N \le \pi_j - \pi_i \le N$, which effectively means there is no constraint on the $\pi$ values.
Note, however, that other edges may still constrain $i$ or $j$ by transitivity.

\paragraph{Fusion-preventing edges}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (fusion-preventing edges from $i$ to $j$)      \\
           \> \ldots
\end{tabbing}
For every fusion-preventing edge, we add the constraint $x_{ij} = 1$, so that no fusion can occur.
This, together with the precedence constraints above, has the effect of enforcing $\pi_i < \pi_j$.

\paragraph{Fusion between different types}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\emptyset \wedge gen_j=\emptyset$)                 \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\{i'\}$) \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_j=\{j'\}$) \\
           \> \ldots
\end{tabbing}
Nodes with different loop sizes cannot be fused together, generally.
If two nodes have different sizes or rates, but are both sub-rates of another rate, they \emph{can} be fused together if they are both fused with their generators.
\TODO{Expand}.

\paragraph{Array contraction}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (for all edges from i)            \\
           \> \ldots \\
\end{tabbing}
An array is fully contracted if all its output edges are fused with it.
Thus, $c_i=0$ only if $\forall j | (i,j) \in E. x_{ij} = 0$.
By minimising $c_i$ in the objective function, we favour solutions that reduce the number of required intermediate arrays.

@@

\subsubsection{An optimised version}
We can simplify the version above and remove many constraints and variables by noting that if there is a fusion-preventing edge between $i$ and $j$, $x_{ij} = 1$, so the precedence constraint can be simplified to just $\pi_i < \pi_j$, and the $x_{ij}$ variable removed.
Likewise with clusters of different, incompatible types.
The contraction variable $c_i$ can be removed, if $i$ has any fusion-preventing output edges: any such edges use the array and cannot be fused, thus there is no possibility of the array being contracted away.
